---
title: "canteen time series"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### read in data
```{r, echo=FALSE}
dat = read.csv("can_even.csv.txt",header = T)
library("tseries")
library("forecast")
library("ggplot2")

w.width <- 12*24*28 # define the length of train data
leap <- 12*24 #freq to renew the model 1 day
long <- msts(dat[,2],c((12*24),(12*24*7)))

mov <- function(ts.full,i,s =1)
{s.time <- (s+(leap)*i/leap/7)
e.time <- (s+(w.width-1+leap*i)/leap/7)
samp <- window(ts.full,start=s.time,end=e.time)
return(samp)}
# plot full period JUN 15 00:00-JUL 15 23:55
attach(dat)
strptime(time,format = "%Y-%m-%d %H:%M:%S")->t1
dat$time <- t1
str(dat)
```

```{r, plot data fully}
canfull <-ggplot(data=dat, aes(x=time, y=counts))+geom_line()+scale_x_datetime(name='Period:Jun.15 Wed 00:00 - Jul.15 Fri 23:55',date_labels ="%d-%H",date_breaks="48 hours",limits = c(as.POSIXct(t1[1]),as.POSIXct(t1[length(t1)])))+theme_bw()+ggtitle("Canteen Device Counts");canfull
jpeg("canteen_full.jpg",width = 1000, height = 500,res = 120)
print (canfull)
dev.off()
detach(dat)
wed<- dat[1:(12*24+1),]
str(wed)
canday <- ggplot(data = wed,aes(x=wed$time, y=wed$counts))+geom_line()+xlab("Jun 15")+ylab("Device Counts")+scale_x_datetime(name='Period:Jun.15 Wed 00:00 - 23:55',labels = date_format("%H",tz="Asia/Kuala_Lumpur"),date_breaks="2 hours")+ggtitle("Canteen Device Counts @ Jun 15")+theme_bw();canday
jpeg("canteen_wed.jpg",width = 600, height = 500,res = 120)
print (canday)
dev.off()

wkend<- dat[(12*24*3+1):(12*24*5+1),]
str(wkend)
canwk <- ggplot(data = wkend,aes(x=wkend$time, y=wkend$counts))+geom_line()+xlab("Jun 18")+ylab("Device Counts")+scale_x_datetime(name='Time',labels = date_format("%H",tz="Asia/Kuala_Lumpur"),date_breaks="2 hours")+ggtitle("Canteen Device Counts @ Jun 18-19 Sat- Sun")+theme_bw();canwk
jpeg("canteen_wkend.jpg",width = 600, height = 500,res = 120)
print (canwk)
dev.off()
```
### using hour and week as period
```{r multi.seasonal, echo=FALSE}
ms.tr <- msts(dat[c(1:w.width),2],c(12,(12*24*7)))
plot.ts(x = ms.tr)
```

### function 

* use fourier and AICc to find optimal K combinations for different frequencies, maximum of freq not exceeding T/2  

* eg: total 4 weeks, the maximum would no larger than 2

```{r fourier K}
find.k <- function(dat.ts,f1.min=1,f1.max,f2.min=1,f2.max,f3.max=NULL)
  {
  if (length(f3.max) < 1) {
aicc.tmp <- NULL
aicc.val <- NULL
for (i in f1.min:f1.max) {
  for (j in f2.min:f2.max) {
    x.fourier <- fourier(dat.ts,c(i,j))
    fitmd <- auto.arima(dat.ts, D = 0, max.P = 0, max.Q = 0, xreg = x.fourier,seasonal=FALSE,max.d=2, ic="aicc", allowdrift=TRUE)
    aicc.tmp <- cbind(i, j , fitmd$aicc ,arimaorder(fitmd)[1],arimaorder(fitmd)[2],arimaorder(fitmd)[3])
    aicc.val <- rbind(aicc.val,aicc.tmp)
  }
}
colnames(aicc.val) <- c("F1_terms","F2_terms","AICc","p","d","q")
aicc.val <- data.frame(aicc.val)
min.aicc <- min(aicc.val$AICc)
min.f <-aicc.val[which(aicc.val$AICc == min.aicc),]
}
  else {
  aicc.tmp <- NULL
  aicc.val <- NULL
for (i in f1.min:f1.max) {
  for (j in f2.min:f2.max) {
    for (k in 1:f3.max){
    x.fourier <- fourier(dat.ts,c(i,j,k))
    fitmd <- auto.arima(dat.ts, D = 0, max.P = 0, max.Q = 0, xreg = x.fourier,seasonal=FALSE,max.d=2, ic="aicc", allowdrift=TRUE)
    aicc.tmp <- cbind(i, j , k , fitmd$aicc)
    aicc.val <- rbind(aicc.val,aicc.tmp)
    }
  }
}
colnames(aicc.val) <- c("F1_terms","F2_terms","F3_terms","AICc")
aicc.val <- data.frame(aicc.val)
min.bic <- min(aicc.val$bic)
min.f <-aicc.val[which(aicc.val$bic == min.bic),]
  }
  return(min.f)
}
```
### Test with hour&week cycle
```{r hour&week}
ms.tr <- msts(dat[c(1:w.width),2],c(12,(12*24*7)))

#find.k(ms.tr,4,6,1,2)
#F1_terms F2_terms     AICc p d q
#       5        2 55222.88 3 0 4
```
### Test with day&week cycle
```{r day&week}
ms.dw <- msts(dat[c(1:(w.width+1)),2],c((12*24),(12*24*7)))
#pt<-proc.time()
#find.k(ms.dw,f1.min = 6,f1.max =10,f2.min = 1,f2.max = 2)
#proc.time() - pt
# time used
#   user  system elapsed 
# 589.86    0.19  591.26
#F1_terms F2_terms     AICc p d q
#      10        2 54646.41 5 1 3
#F1_terms F2_terms     AICc p d q
#      15        1 54579.08 2 1 4
#F1_terms F2_terms     AICc p d q
#      25        1 54523.88 3 1 4
```
### using fft to find optimal freq
```{r fft, echo= F}
find.freq <- function(f.ts){
abs(fft(f.ts)/sqrt(length(f.ts)))^2-> I
Pow = I[1:as.integer(length(f.ts)/2)]
FREQ =(0:as.integer((length(Pow)-1)))/length(Pow)
plot(FREQ,Pow,type="l")
T1 <- data.frame(array(c(Pow,FREQ),dim=c(length(Pow),2)))
colnames(T1)<- c("Power","Frequency")

return (T1)
}
t1<-find.freq(dat[c(1:w.width),2]) 
pow <- order(t1$Power,decreasing = T) 
# power highest  T=Inf/144/1008
ff1 <- 1/t1$Frequency[pow[2]]
ff1
## power has T=24/28
t2<- find.freq(diff(dat[c(1:w.width),2]))
pow <- order(t2$Power,decreasing = T) 
# power highest T=28.79/24
ff1 <- 1/t2$Frequency[pow[1]]
ff1
```

### Produce forecasting and model check use the Optimal K
```{r forecasting,echo=F}

build.fmd <- function(dat.ts,f1.k,f2.k=1,len.fcast = leap){
x.fd = fourier(dat.ts,c(f1.k,f2.k))
#forecasting following time
x.fdh = fourier(dat.ts,c(f1.k,f2.k), h = len.fcast) 
best_md.day = auto.arima(dat.ts, D = 0, xreg = x.fd, max.d=3, ic="bic",seasonal = F, allowdrift=TRUE)
b_dfcast = forecast(best_md.day, xreg = x.fdh, h = leap)
full <- list("bst_fcast" =b_dfcast,"fittedmd" = best_md.day)
return(full)}


proc.time()-pt
```

### display results of forecast with best model
```{r display forecast}

fcast.ori.display <- function(forecast, ori,...){
  # data wrangling
  time <- attr(forecast$x, "tsp")
  time <- seq(time[1], attr(forecast$mean, "tsp")[2], by=1/time[3])
  lenx <- length(forecast$x)
  lenmn <- length(forecast$mean)
 
  df <- data.frame(time=time,
    x=c(forecast$x, forecast$mean),
    forecast=c(rep(NA, lenx), forecast$mean),
    orival = c(rep(NA, lenx), ori),
    low1=c(rep(NA, lenx), forecast$lower[, 1]),
    upp1=c(rep(NA, lenx), forecast$upper[, 1]),
    low2=c(rep(NA, lenx), forecast$lower[, 2]),
    upp2=c(rep(NA, lenx), forecast$upper[, 2])
  )
 
p<-  ggplot(df, aes(time, x)) +
    geom_linerange(aes(ymin=low2, ymax=upp2), color="grey", na.rm=TRUE) +
    geom_linerange(aes(ymin=low1, ymax=upp1), color="lightblue", na.rm=TRUE) +
    geom_line() +
    geom_line(data=df[!is.na(df$forecast), ], aes(time, forecast), color="blue", na.rm=TRUE) +geom_line(data=df[!is.na(df$orival), ], aes(time, orival), color="red",linetype = 2, na.rm=TRUE) 
figf <-p+ggtitle(paste("Forecasts from ", forecast$method))
return(figf)}
## only plot the predictions
fc.ori.con <- function(forecast, ori,...){
  # data wrangling
  time <- attr(forecast$mean, "tsp")
  time <- seq(time[1], attr(forecast$mean, "tsp")[2], by=1/time[3])
  lenx <- length(forecast$x)
  lenmn <- length(forecast$mean)
 
  df <- data.frame(time=time,
    x=c(forecast$mean),
    forecast=c( forecast$mean),
    orival = c( ori),
    low1=c(forecast$lower[, 1]),
    upp1=c(forecast$upper[, 1]),
    low2=c(forecast$lower[, 2]),
    upp2=c( forecast$upper[, 2])
  )
 
p<-  ggplot(df, aes(time, x)) +
    geom_linerange(aes(ymin=low2, ymax=upp2), color="grey", na.rm=TRUE) +
    geom_linerange(aes(ymin=low1, ymax=upp1), color="lightblue", na.rm=TRUE) +
    geom_line() +
    geom_line(data=df[!is.na(df$forecast), ], aes(time, forecast), color="blue", na.rm=TRUE) +geom_line(data=df[!is.na(df$orival), ], aes(time, orival), color="red",linetype = 2, na.rm=TRUE) 
figf <-p+ggtitle(paste("Forecasts", forecast$method))
return(figf)}
```

```{r start plotting comparison}
dw.md <- build.fmd(ms.dw,15,6)
tst <-dat[c((w.width+1):(w.width+leap)),2]
# 15,6 1st MAE 12.99 (3,1,4)
#fd.fig <- fcast.ori.display(dw.md$bst_fcast,ori = tst)
f.fig <-fc.ori.con(dw.md$bst_fcast,ori = tst)
jpeg("1stday.jpg",width = 600,height = 500,bg="transparent",res=120)
f.fig
dev.off()

dw3.md <-build.fmd(ms.dw3,10,2)
#10,2,MAE17.92
#15,6,MAE17.51
jpeg("3day15_6.jpg",width = 600,height = 500,bg="transparent",res=120)
fc.ori.con(dw3.md$bst_fcast,ori = tst3)
dev.off()
print (paste("mean error:",sum(abs(c(dw3.md$bst_fcast[["mean"]])-tst3))/length(tst3)))
```

#### change the time span and explore
```{r day&week with new period}
pt <- proc.time()
ms.tr <- msts(dat[(c(1:w.width)+leap),2],c(12*24,(12*24*7)))
tsdisplay(x = ms.tr)
tst.tr <- dat[c((w.width+1+leap):(w.width+leap+leap)),2]
#kmin<-find.k(ms.tr,f1.min = 10,f1.max = 15,f2.min = 1,f2.max = 2)
#print(kmin)

ff1 <- build.fmd(ms.dw2,10,2)
p2 <- fc.ori.con(ff1$bst_fcast,tst.tr)
# 10,2 2nd MAE 15.82, (4,1,4)
# 15,6 MAE = 12.88
ff3 <-build.fmd(ms.dw3)
residual <- tb.ff$mean - tst.tr
fd.fig <-fc.ori.con(tb.ff,ori = tst)
jpeg("1stdayTBATS.jpg",width = 600,height = 500,bg="transparent",res=120)
fd.fig
dev.off()

tb.fd2 <-tbats(ms.dw2)
tb.ff2 <-forecast(tb.fd2,h=12*24)
tp2<- fc.ori.con(tb.ff2,tst.tr)
jpeg("2nddayTBATS.jpg",width = 600,height = 500,bg="transparent",res=120)
tp2
dev.off()


tst3 <- dat[c((w.width+1+leap*2):(w.width+leap*3)),2]
ms.dw3 <- mov(long,2)
tb.md3<- tbats(ms.dw3)
tb.ff3<- forecast(tb.md3,h=leap)
jpeg("3rddayTBATS.jpg",width = 600,height = 500,bg="transparent",res=120)
fc.ori.con(tb.ff3,tst3)
dev.off()

print (paste("mean error:",sum(abs(c(tb.ff3[["mean"]])-tst3))/length(tst3)))

res3<- (tb.ff3$mean-tst3)/tst3
res3a<- (dw3.md$bst_fcast[["mean"]]-tst3)/tst3
plot(res3a,type ="n",main= paste("Error Rate(R:ARIMA-B:TBATS) "),ylab="")
lines(res3a,col="red")
lines(res3,col="blue")
lines(h=mean(abs(res3a)),lty=1,col="red")
abline(h=mean(abs(res3)),lty=1,col="blue")

res1<- (tb.ff$mean-tst)/tst
res1a<- (ff$bst_fcast[["mean"]]-tst)/tst
plot(res1a,type ="n",main= paste("Error Rate(R:ARIMA-B:TBATS) "),ylab="")
lines(res1a,col="red")
lines(res1,col="blue")
lines(h=mean(abs(res3a)),lty=1,col="red")
abline(h=mean(abs(res3)),lty=1,col="blue")


#MAE & MAPE
print (paste("mean error:",sum(abs(c(tb.fc$mean)-tst)/tst)/length(tst)))
plot((c(ff$bst_fcast[["mean"]])-tst3))

print (paste("mean error rate:" ,sum(abs(tb.ff3[["mean"]]-tst3)/tst3)/length(tst3)))

#TBATTS: 0.367

#tsdisplay(best_md.day$residuals)                   
print(paste("used time",proc.time()-pt))
```


### dynamic linear models with DLM
```{r Dlm}
### DLM mod test 
library(dlm)
build.SDE<- function(x) {
dlmModSeas(12*24*7) +dlmModPoly() ->mod
  V(mod)<- exp(x[1])
  diag(W(mod))[1:3]<-exp(x[2:4])
  return(mod)}
#+dlmModPoly(1, dV = x[1], dW = x[2]) }

fit.SDE <- dlmMLE(ms.dw, parm = rep(0, 4), build = build.SDE,lower = rep(1e-8, 2), hessian = TRUE)
unlist(fit.SDE$par)

```

